General constraints on magic distillation protocols from statistical mechanics

### 2021-06-04 magic_paper.tex

\begin{definition} A continuous, real-valued function $f$ defined on $\D \subseteq \mathbb{R}^N$ is called Schur-convex on the domain $\D$ if $\p \prec \q$ on $\D$ implies that $f(\p) \le f(\q)$. A continuous function $f:\D \rightarrow \mathbb{R}$ is called Schur-concave if $(-f)$ is Schur-convex.
\end{definition}
We can consider the set of all quasi-distributions, which form a hyper-plane and are given by
\begin{equation}
\D = \{ \w \in \mathbb{R}^N: \sum_{i=1}^N w_i = 1\}.
\end{equation}
For functions that are symmetric, namely $f(\p) = f(\Pi \p)$ for any component permutation $\Pi$, it suffices to restrict to the following region,
\begin{equation}
\D^\downarrow = \{ \w \in \D : w_1 \ge w_2 \ge \cdots \ge w_N\}.
\end{equation}

We now have the following theorem~\cite{cit:marshall} that provides a characterization of symmetric, Schur-convex and Schur-concave functions on quasi-distributions.
\begin{lemma} 
Let $f$ be a real-valued continuous function defined on $\D^\downarrow$ that is continuously differentiable on the interior of $\D^\downarrow$. Then, $f$ is Schur-convex on $\D^\downarrow$ if and only if $\partial_k f(\w)$ is decreasing in $k$ for all $\w \in \D^\downarrow$, namely
\begin{equation}
\partial_1 f(\w) \ge \partial_2 f(\w) \ge \cdots \ge \partial_N f(\w),
\end{equation}
for all $\w$ in the interior of $\D^\downarrow$.
\end{lemma}

@@

\begin{theorem}
	Let $\w, \w' \in \mathbb{R}^N$ be quasi-distributions such that $\w \succ \w'$.
	Then, $H_\alpha(\w) \leq H_\alpha(\w')$, where $\alpha$ is such that $\hspace{-1pt}H_\alpha(\x)\hspace{-2pt}$ is well-defined for all quasi-distributions $\x$.
\end{theorem}
\begin{proof}
Given a fixed $\alpha$ for which $H_\alpha(\x)$ is well-defined for all quasi-distributions $\x$, the function $H_\alpha$ is real-valued, smooth on the interior of $\D^\downarrow$ and symmetric on permutation components of $\x$.

Consider the partial derivatives
\begin{equation}
	\frac{\partial H_\alpha}{\partial w_i} = \frac{\alpha w_i^{\alpha-1}}{(1-\alpha)\sum_k{w_k^\alpha}} 
\end{equation}
Whenever $w_i \geq w_j$, we have that 
\begin{equation}
	\frac{w_i^{\alpha-1}}{1-\alpha} \leq \frac{w_j^{\alpha-1}}{1-\alpha},
\end{equation}
independently of whether $\alpha < 1$ or $\alpha > 1$, which leads to
\begin{equation}
	\frac{\partial H_\alpha}{\partial w_i} \leq \frac{\partial H_\alpha}{\partial w_j}.
\end{equation}
Therefore, $H_\alpha$ is Schur-concave and
\begin{equation}
	H_\alpha(\w) \leq H_\alpha(\w'),
\end{equation}
for any quasi-distributions $\w, \w'$.
\end{proof}

@@

>>>>

\begin{definition} 
A continuous, real-valued function $f$ defined on $\mathcal{Q} \subseteq \mathbb{R}^N$ is called Schur-convex on the domain $\mathcal{Q}$ if $\p \prec \q$ on $\mathcal{Q}$ implies that $f(\p) \le f(\q)$. A continuous function $f:\mathcal{Q} \rightarrow \mathbb{R}$ is called Schur-concave if $(-f)$ is Schur-convex.
\end{definition}
We can consider the set of all quasi-distributions, which form a hyper-plane and are given by
\begin{equation}
\mathcal{Q} = \left\{ \w \in \mathbb{R}^N: \sum_{i=1}^N w_i = 1 \right\}.
\end{equation}
For functions that are symmetric, namely $f(\p) = f(\Pi \p)$ for any component permutation $\Pi$, it suffices to restrict to the following region,
\begin{equation}
\mathcal{Q}^\downarrow = \{ \w \in \mathcal{Q} : w_1 \ge w_2 \ge \cdots \ge w_N\}.
\end{equation}

We now have the following theorem~\cite{cit:marshall} that provides a characterization of symmetric, Schur-convex and Schur-concave functions on quasi-distributions.
\begin{lemma}
Let $f$ be a real-valued continuous function defined on $\mathcal{Q}^\downarrow$ that is continuously differentiable on the interior of $\mathcal{Q}^\downarrow$. 
Then, $f$ is Schur-convex on $\mathcal{Q}^\downarrow$ if and only if $\partial_k f(\w)$ is decreasing in $k$ for all $\w \in \mathcal{Q}^\downarrow$, namely
\begin{equation}
\partial_1 f(\w) \ge \partial_2 f(\w) \ge \cdots \ge \partial_N f(\w),
\end{equation}
for all $\w$ in the interior of $\mathcal{Q}^\downarrow$.
\end{lemma}

@@

\begin{theorem}
	Let $\w, \w' \in \mathbb{R}^N$ be quasi-distributions such that $\w \succ \w'$.
	Then, $H_\alpha(\w) \leq H_\alpha(\w')$, where $\alpha$ is such that $\hspace{-1pt}H_\alpha(\x)\hspace{-2pt}$ is well-defined for all quasi-distributions $\x$.
\end{theorem}
\begin{proof}
Given a fixed $\alpha$ for which $H_\alpha(\x)$ is well-defined for all quasi-distributions $\x$, the function $H_\alpha$ is real-valued, smooth on the interior of $\mathcal{Q}^\downarrow$ and symmetric on permutation components of $\x$.

Consider the partial derivatives
\begin{equation}
	\frac{\partial H_\alpha}{\partial w_i} = \frac{\alpha w_i^{\alpha-1}}{(1-\alpha)\sum_k{w_k^\alpha}} 
\end{equation}
Whenever $w_i \geq w_j$, we have that 
\begin{equation}
	\frac{w_i^{\alpha-1}}{1-\alpha} \leq \frac{w_j^{\alpha-1}}{1-\alpha},
\end{equation}
\nick{wrong} independently of whether $\alpha < 1$ or $\alpha > 1$, which leads to
\begin{equation}
	\frac{\partial H_\alpha}{\partial w_i} \leq \frac{\partial H_\alpha}{\partial w_j}.
\end{equation}
Therefore, $H_\alpha$ is Schur-concave and
\begin{equation}
	H_\alpha(\w) \leq H_\alpha(\w'),
\end{equation}
for any quasi-distributions $\w, \w'$.
\end{proof}

@@

\ddd{[TO-DO AND TIDY AND CHECK]} The extension of the domain of the Renyi entropies to real values means that is it not real-valued for certain values of $\alpha$. However we can make $H_\alpha(\w)$ well-defined on the whole of $\mathcal{Q}$ if $\alpha = \frac{a}{b}$ where $b$ is an odd integer, and $a$ is an even integer. Depending on the set of quasi-distributions the quantum system supports the restriction of $a$ being even could be lifted, provided that the sum $\sum_i w_i^\alpha >0$, however we do not consider this here, and restrict to the assumption that $a$ is even.

\begin{theorem} The functions $H_\alpha(\w)$ are well-defined on $\mathcal{Q}$ for $\alpha = \frac{a}{b}$ with $a$ an even integer and $b$ an odd integer, and are Schur-concave functions.
\end{theorem}
\begin{proof}
We now make use of the Schur-Ostrowski criterion to determine when $H_\alpha$ is Schur-Convex on $\mathcal{Q}$.

We have that
\begin{align}
V_k:=-\frac{\partial H_\alpha}{\partial w_k} &= \partial_k \frac{1}{\alpha-1} \log \sum_i w_i^{\alpha} \\
&= \frac{1}{(\alpha-1) \sum_i w_i^{\alpha}}\alpha w_k^{\alpha-1}.
\end{align}
We now impose the monotonicity condition $V_1 \ge V_2 \ge V_3 \cdots \ge V_N$. This requires, setting $S= \sum_i w_i^\alpha$ for simplicity, that
\begin{align}
 \frac{1}{(\alpha-1)S}\alpha w_i^{\alpha-1}\ge \frac{1}{(\alpha-1)S}\alpha w_{i+1}^{\alpha-1},
\end{align}
for each $i=1,2,\dots, N-1$. We have that $S >0$, since $a$ is even, and assume that $\alpha >1$ so that
\begin{equation}
w_i^{\alpha-1}\ge w_{i+1}^{\alpha-1}\mbox{ for all } 1 \le i < N.
\end{equation}
For $\alpha = a/b > 1$ we have $\alpha-1 = (a-b)/b = a'/b >0$ where $a'$ is an even integer and $b$ an odd integer. We now assume that $\w \in \mathcal{Q}^\downarrow$ so that $ w_i \ge w_{i+1}$. \ddd{[Wait I'm not sure the inequalities work...]}
\end{proof}
